### 2024-09-12 tables-prev/tma-2023/errata-stats-stream.tex
Initial import of text on documents from IMC 2021 and TMA 2023 papers

\begin{table*}
\centering
\footnotesize
\begin{tabular}{lr|rrrr|rr}
\toprule
\textbf{Stream} & \textbf{\#} & \textbf{Verified} & \textbf{Held} & \textbf{Rejected} & \textbf{Reported} & \textbf{Technical} & \textbf{Editorial} \\
\midrule
IETF (6619) & 5797  & 2348  & 1815  & 798  & 836  & 3034  & 2763 \\
IAB (124) & 55  & 25  & 13  & 3  & 14  & 23  & 32 \\
Independent (376) & 330  & 235  & 32  & 28  & 35  & 172  & 158 \\
Legacy (1929) & 510  & 226  & 182  & 39  & 63  & 198  & 312 \\
IRTF (97) & 67  & 37  & 8  & 9  & 13  & 44  & 23 \\
\midrule All & 6759  & 2871  & 2050  & 877  & 961  & 3471  & 3288 \\
\bottomrule
\end{tabular}
\caption{Errata statistics by stream; the ``\emph{Editorial}'' stream has no documents, and is not shown.}
\label{tab:errata-stats-stream}
\end{table*}