### 2024-09-12 tables-prev/tma-2023/top-rfcs-errata.tex

\begin{table*}
\centering
\footnotesize
\begin{tabular}{rlrrr}
\toprule
\textbf{RFC} & \textbf{Title} & \textbf{Year} & \textbf{Area} & \textbf{Filing count} \\
\midrule
4601 & Protocol Independent Multicast - Sparse Mode (PIM-SM): Protocol Specification (Revised) & 2006 & rtg & 114 \\
4880 & OpenPGP Message Format & 2007 & sec & 52 \\
793 & Transmission Control Protocol & 1981 & None & 47 \\
4634 & US Secure Hash Algorithms (SHA and HMAC-SHA) & 2006 & None & 44 \\
5661 & Network File System (NFS) Version 4 Minor Version 1 Protocol & 2010 & tsv & 42 \\
1345 & Character Mnemonics and Character Sets & 1992 & app & 41 \\
8446 & The Transport Layer Security (TLS) Protocol Version 1.3 & 2018 & sec & 40 \\
5545 & Internet Calendaring and Scheduling Core Object Specification (iCalendar) & 2009 & app & 35 \\
3261 & SIP: Session Initiation Protocol & 2002 & rai & 33 \\
5905 & Network Time Protocol Version 4: Protocol and Algorithms Specification & 2010 & int & 32 \\
\bottomrule
\end{tabular}
\caption{Top 10 RFCs by errata filing count}
\label{tab:top-10-rfcs-by-errata}
\end{table*}